---
title: "Untitled"
author: "Li Ying"
date: "29 June 2015"
output: html_document
---

```{r}
library(xlsx); library(hash); library(dplyr); library(ggplot2); library(corrplot)
# Load data
setwd("~/Dropbox/#0")
word<-read.xlsx("./data.xlsx",
sheetIndex=4,header=FALSE,colIndex=2:1000,rowIndex=4:76)
cues<-1:73

word<-data.frame(cue=cues,word)



```


```{r}
# collapse data
collapse<-apply(word[,-1],1,paste,collapse=" ")
collapse<-data.frame(cue=word$cue,Assoc=collapse)

# Transform text column to factor
collapse$Assoc<-as.character(collapse$Assoc)
dim(collapse) #[1] 73  2
# Delete NA participants; Nothing changed
collapse<-na.omit(collapse)
dim(collapse) #[1] 73  2

# Clean
require("tm")
text_corpus<-Corpus(VectorSource(collapse$Assoc))
clean<-tm_map(text_corpus,tolower)
clean<-tm_map(clean,removeNumbers)
clean<-tm_map(clean,removeWords,stopwords())
clean<-tm_map(clean,stripWhitespace)
clean<-tm_map(clean,PlainTextDocument)

dtm <- DocumentTermMatrix(clean)
dim(dtm) #[1]   73 1780
uniwords<-dtm$dimnames$Terms
length(unique(uniwords))
```

```{r}
# reduce dataset 
freq_dict<-findFreqTerms(dtm,1)
freq <- DocumentTermMatrix(clean,list(dictionary=freq_dict))
colnames<-freq$dimnames$Terms

freq<-matrix(freq,nrow= length(cues),ncol=length(colnames),dimnames=list(cues,colnames) )
freq<-t(freq)
```

### Plot histogram
```{r}
freqt<-t(freq)
freqDF<-data.frame(freqt)
freqDF<-data.frame(cue=1:73,freqDF)
freqDF<-mutate(freqDF, response_number=apply(freqDF[,2:1781],1,sum))
freqDF$response_number
freqDF<-freqDF[,c("cue","response_number")]
dev.off()
par(mar=c(4,4,2,2))
par(mfrow=c(1,2))
hist(freqDF$response_number,main="Histogram of response number", xlab="No. of Associations", ylab="frequency")
qqnorm(freqDF$response_number, col="blue") ;qqline(freqDF$response_number, col=2)

```

### Association Matrix
```{r}
corTable<-cor(freq)
dev.off()
par(mar=c(1,2,2,2))
corrplot(cor(freq, use="pairwise.complete.obs"), method="shade",shade.col=NA, tl.col="black", tl.cex=0.4, tl.srt=90)  
```

### Planar Maximally Filtered Graph [***PMGF***]
- The complete association correlation network for N target words contains N (N {1) edges.
- Since most of the edges have small values (weak correlations), the relevant information about the network (e.g. topology, organization), can be obscured.
- To overcome this obstacle, use complete network to construct a sub-graph that captures the most relevant information embedded in the original network. 
- How to do that? `PMGF` -- an expansion from MST (Minimum Spanning Tree)

#### Consrtuction of Planr Maximally Filtered Graph 
- 1. order the N (N - 1) values of the correlation matrix C in decreasing rank.
- 2. start from pair of nodes with highest correlational value, say i and j, with the highest correlation and draw a link i --> j between them.
- 3. The process continues according to the rank order where in each iteration a link is added ***if and only if*** the resulting graph (network) is still planar, i.e. it can be drawn on the surface of a sphere without link crossing

Learn this: http://www.r-bloggers.com/network-visualization-in-r-with-the-igraph-package/

### Minimum Spanning Trees (MST)
- It is a edge-weighted graph where edge length is associated with dissimilarity between variables.
- dissimilarity is computed by [Bray–Curtis dissimilarity](https://en.wikipedia.org/wiki/Bray%E2%80%93Curtis_dissimilarity):a statistic used to quantify the compositional dissimilarity between two different sites, based on counts at each site.


[Detailed Algorithm on MST](http://algs4.cs.princeton.edu/43mst/)
```{r Plot Minimum Spanning Trees}
library(vegan);
data(dune) # 30 species on 20 cite
dis <- vegdist(t(freq))
tr <- spantree(dis)
## Add tree to a metric scaling 
plot(tr, cmdscale(dis), type = "t")
## Find a configuration to display the tree neatly
plot(tr, type = "t")
```

### Plot graph. Association matrix can only be plotted on sqaure matrix
Method: [Tutorial](http://stackoverflow.com/questions/19965600/igraph-r-how-to-create-correlation-network-with-only-strong-r-values)
```{r}
library(igraph)

# build a graph from the above matrix
diag(corTable) <- 0
g <- graph.adjacency(corTable, weighted=T, mode = "undirected")
g <- delete.edges(g, E(g)[ weight < 0.2 ])
#freq# remove loops
g <- simplify(g)
# set labels and degrees of vertices
V(g)$label <- V(g)$name
V(g)$degree <- degree(g)
par(mar=c(2,2,2,2))
plot(g,vertex.size=3,vertex.label=NA)
```


### Calculate parameters
```{r}
average.path.length(g)
diameter(g)
transitivity(g)
mean(degree(g))
par(mar=c(4,4,4,4))
hist(degree(g),col="steelblue", main="Degree Histogram", xlab="degree/number of edge")
plot(degree.distribution(g))
```


### Note
[Degree](http://mathinsight.org/degree_distribution)
[Unweighted Adjacency Matrix](http://mathinsight.org/definition/adjacency_matrix)
[Small Word Network](http://mathinsight.org/small_world_network): Featured by large clustering coefficient and small average path length 
[Transitivity](http://mathinsight.org/evidence_additional_structure_real_networks#transitivity)
[Transitivity math](http://mathinsight.org/definition/transitivity_graph)
[Clustering Coefficient](http://mathinsight.org/definition/clustering_coefficient)

[Free Will Scale Scoring Document](http://agencyandresponsibility.typepad.com/files/competing-free-will-and-responsibility-scales.pdf)
[Minimum Spanning Tree Tutorial](http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf)

