---
title: "iGraph Tutorial"
author: "Li Ying"
date: "1 July 2015"
output: html_document
---
##2.1 Creating graphs

- The objects we study in this course are graphs (or networks).  
- They consist of a set of nodes and a set of edges.  
- In this command, we are assigning to the variable g a graph that has nodes V = {1,2,3,4,5} and has edges E = {(1,2),(1,3),(2,3),(3,5)}
```{r}
library(igraph)
g <- graph( c(1,2, 1,3, 2,3, 3,5), n=5 )
V(g)
E(g)

```

#### Add nodes and edges to an already existing graph, e.g.:

```{r}
g <- graph.empty() + vertices(letters[1:10], color="red"); plot(g)
g <- g + vertices(letters[11:20], color="blue"); plot(g)
g <- g + edges(sample(V(g), 30, replace=TRUE), color="green");plot(g)
V(g)


```

###A bug: can't start with 0
```{r}
karate <- read.graph("http://cneurocvs.rmki.kfki.hu/igraph/karate.net", format="pajek")
V(g); vcount(g)
E(g); ecount(g)
```

### Manipulating attributes in graphs
```{r}
g <- erdos.renyi.game(8, 0.5); plot(g) # generate a random graph with 10 nodes, assigns
V(g)$color <- sample( c("red", "black"), vcount(g), rep=TRUE); plot(g) # Assign random color to node
E(g)$color <- "grey"; plot(g) # Color all edge in grey
red <- V(g)[ color == "red" ]
bl <- V(g)[ color == "black" ]
E(g)[ red %--% red ]$color <- "red"; plot(g) 
E(g)[ bl %--% bl ]$color <- "black"; plot(g) 


#The next example assigns random weights to a lattice graph and then colors the ones having weight over 0.9 red, and the rest grey.
dev.off()
g <- graph.lattice( c(10,10) )
E(g)$weight <- runif(ecount(g))
E(g)$color <- "grey"
E(g)[ weight > 0.9 ]$color <- "red"
plot(g)

# a more refined version
dev.off()
g <- graph.lattice( c(10,10) )
E(g)$weight <- runif(ecount(g))
E(g)$color <- "grey"
E(g)[ weight > 0.9 ]$color <- "red"
plot(g, vertex.size=2, vertex.label=NA, layout=layout.kamada.kawai,
edge.width=2+3*E(g)$weight)

```

### Visualizing graphs
A very important part in the analysis of networks is being able to visualize them. As an example the following commands render the three graphs depicted in the figure below.
```{r}
er_graph <- erdos.renyi.game(100, 2/100)
plot(er_graph, vertex.label=NA, vertex.size=3) # change size and lable #Erd ̈os-R ́enyi
ws_graph <- watts.strogatz.game(1, 100, 4, 0.05)
plot(ws_graph, layout=layout.circle, vertex.label=NA, vertex.size=3) #Watts-Strogatz
ba_graph <- barabasi.game(100)
plot(ba_graph, vertex.label=NA, vertex.size=3) #Barab ́asi-Albert

```

The plot command is very flexible and has many parameters that control the behavior of the visualization. For greater details: [click](http://127.0.0.1:28744/library/igraph/html/plot.common.html)

### Measuring graph
- `average.path.length`
- `diameter`
- `clustering coefficient` or `transitivity`
```{r}
g <- graph.lattice( length=100, dim=1, nei=4 )
average.path.length(g)
diameter(g)

ws <- watts.strogatz.game(1, 100, 4, 0.05)
plot(ws)
transitivity(ws)
degree(ws)
hist(degree(ws))
plot(degree.distribution(ws))
```



